A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-02-07 paper.tex
Update on Overleaf.

\glspl{iam} are tools used to understand the complex interactions between
energy, economy, land use, water, and climate systems. \glspl{iam} provide
global projections of systemic change by dividing the world into a number of
representative regions (typically 10 to 30), the definition of which is distinct
for each model \cite{krey_global_2014}. Results from \glspl{iam} are integral in
a number of international studies, which notably include projections of climate
and economic futures. Recently, the \gls{iam} community has developed scenarios
based on the \glspl{ssp} \cite{fujimori_ssp3:_2017, van_vuuren_energy_2017,
  calvin_ssp4:_2017, fricko_marker_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The \glspl{ssp} are designed to
be used in research that include \gls{esm} simulations, climate impact,
adaptation and climate mitigation studies \cite{vuuren_new_2013}.

@@

Separately, if there are large discrepancies between the model results in the
base year and the historic data used for harmonization, convergence methods can
result in harmonized trajectories that do not faithfully represent the
underlying drivers of emissions. Furthermore, if models report negative
emissions, as is possible in scenarios designed to depict the deployment of
climate mitigation policies with large \cotwo reductions and eventual storage,
then end-of-century emissions characteristics can be of chief import in order to
faithfully match the design parameters of the original scenarios, such as global
mean temperature and other climate metrics.

@@

A number of the classic methods are implemented in \code{aneris} including
ratio-convergence shown in Equation \ref{eqs:ratio}, offset-convergence shown in
Equation \ref{eqs:offset}, and interpolation shown in Equation
\ref{eqs:interpolate}. In all equations the region, species, and sector indices
have been dropped for clarity. Each equation is a function of time, model
trajectory, historical trajectory, base year ($t_i$), and a convergence year
($t_f$), at which point the harmonized trajectory converges to the unharmonized
trajectory. \codeb{aneris} provides a number of methods to choose from for each
of the harmonization families. A summary of all available methods is provided in
Table \ref{tab:meths}.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{decision_tree.pdf}
    \caption[]{
      \label{fig:decision_tree}
      The default method decision tree used in the \code{aneris} software
      library. For all decisions, upper (purple) branches represent a ``yes''
      response and lower (orange) branches represent a ``no'' response. The
      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
      decision-making thresholds for $c_v$ and $dH$ are described below.  Where
      present, convergence years of default methods are provided below the
      method name in parentheses.  Convergence years are chosen in order to
      balance the three harmonization goals discussed in Section
      \ref{sec:concept}. Methods labeled in \textit{green} are likely to closely
      match unharmonized results, methods in \textit{yellow} will likely
      somewhat match unharmonized results, and methods in \textit{red} can be
      expected to have a large relative difference between harmonized and
      unharmonized results.}
  \end{center}
\end{figure}

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is rare but possible for models to report zero
values in the harmonization year in situations in which technologies are
introduced in future time periods in regions or for sectors which produce an
emissions species that is absent in the initial modeling period. In such cases,
an offset method is required as a ratio method would mask future emissions and
erroneously harmonize the trajectory.

In most cases, however, models do report values in the harmonization
year. Figure \ref{fig:cases} displays a number of example trajectories which
highlight the possible issues resulting from harmonizing model results in
different contexts. These trajectories do not correspond to specific model
results; rather they serve as illustrative examples of the kinds of trajectories
observed in practice. Panels \code{a} and \code{b} highlight examples where
model results peak mid-century, behavior that is seen in a number of scenarios
with general emissions mitigation effects, such as pollution controls applied by
developing nations on transport and industry sectors. Panel \code{a} highlights
a case where model base-year values and history are relative close whereas Panel
\code{b} shows a situation where model values and history are relatively far
apart. Panels \code{c} and \code{d} show similar model trajectories that peak
mid-century but also have negative emissions. Models can report negative
emissions in contexts in which a gas species which can be extracted from the
environment and stored, as is the case for \cotwo in future scenarios with
climate mitigation enacted via the deployment of carbon capture and
sequestration technologies or direct carbon capture technologies. Again, the
relative difference with historical values differ between the panels to explore
harmonization method choices in each situation.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
\gls{iam} values used in the \gls{ssp} and (ongoing) \gls{cmip6} inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

@@

The \texttt{Harmonizer} is a class whose concern is to harmonize model value
trajectories given historical data and possible user method \textit{overrides},
i.e., non-default methods (described further in Section \ref{sec:overrides}). It
is used by the \texttt{HarmonizationDriver}; however it is also available to the
user as a first-class object. The \texttt{Harmonizer} requires that input data
conform to the \code{aneris} calculation data format, which explicitly separates
the emissions species from the sector contributing the emissions (these are
combined in the single \code{Variable} column in the standard \gls{iam}
format). Because the \texttt{Harmonizer} is designed to operate on a single
instance of a model and scenario, the canonical data format includes
\texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
without extraneous meta-data columns for the model and scenario. Once configured
with appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
    \caption[]{
      \label{fig:software}
      The various objects and their relation to one another in the \code{aneris}
      code base as well as a short description of their scope of concern.  
    }
  \end{center}
\end{figure}

@@

%% \begin{figure}
%%   \subfloat[Python Implementation\label{fig:software}]{
%%     \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
%%     }
%%   \subfloat[Workflow\label{fig:workflow}]{
%%     \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
%%     }
%%     \caption{ Panel \ref{fig:software} shows the various objects in the
%%       \code{aneris} code base as well as a short description of their scope of
%%       concern. Panel \ref{fig:workflow} displays the full harmonization process
%%       as executed by \code{aneris} for a single instance of a model and
%%       scenario. Operations that can be configured with user-based input
%%       configurations are shown in purple. Operations governed by the
%%       \texttt{HarmonizationDriver} are shown in red. Data processing operations
%%       are shown in blue. The core harmonization process, governed by the
%%       \texttt{Harmonizer} is shown in green.  }
%%     \label{fig:software-workflow}
%% \end{figure}

The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
by cleaning input data. Cleaning operations include adding (0-valued) model
trajectories that exist in the historical data set but are not provided by the
model input and detecting any issues that would cause the harmonization process
to fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
spatial aggregation to common analysis regions is performed. For example, the
5-region aggregation developed in the \glspl{rcp}
\cite{vuuren_representative_2011} process is commonly used in the \gls{iam}
community and is shown in Figure \ref{fig:regions}. Finally, any exogenous
trajectories the user provides are added. Exogenous trajectories are normally
provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
chlorofluorocarbons provided by the \gls{wmo} \cite{wmo2014}. Upon completion,
the harmonized trajectories and meta data regarding the harmonization process
are returned. A description of all returned meta data is provided in Table
\ref{tab:metadata}.

@@

In order to help identify cases where overrides may be of interest,
harmonization \textit{diagnostics} are provided which analyze the relative
difference between harmonized and unharmonized trajectories at their mid and
end-points. These values are configurable by the user, but defaults of $400$\%
and $200$\%, respectively, are provided based on experiences of the authors'
use of \code{aneris} to date.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from
the \gls{ssp} scenario library\footnote{We refer the reader to the broad
  literature discussing the SSPs, e.g.,
  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
  depth discussion of the scenario architecture and design.} are presented. The
SSP2 \cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
(referred to as SSP2-Ref) is chosen to be analyzed because MESSAGE-GLOBIOM is
the marker scenario\footnote{The concept of ``marker'' scenarios are used to
  illustrate distinctive scenario types within the SSP scenario framework. See
  \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
  description.} for this \gls{ssp}. This SSP2 scenario lies between two
\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing is a
  measurement of the capacity for different emissions species to affect
  tropopausal irradiance and has a relationship to global mean equilibrium
  temperature change. We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
  for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
additionally present the results for the SSP2-based mitigation scenario leading
to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
scenario is chosen because mitigation technologies and policies are enacted
causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
negative \cotwo emissions in some regions and sectors due to carbon capture and
sequestration and afforestation. A scenario in which negative emissions play a
role in mitigation strategies is particularly important because of the
sensitivity of key indicators, such as end-of-century radiative forcing (which
is used to estimate mean global temperature response), to the timing and
magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
scenarios represent two bounding cases in the use of a harmonization approach
and thus provide a case study as to its general applicability. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
aggregate \gls{ghg} emissions, in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the \glspl{rcp}. Historical
data is taken from previously described \gls{luc} and anthropogenic sources,
which comprise 10 separate pollutant and \gls{ghg} species and 12 sectors shown
in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
each scenario, and therefore 1940 trajectories were harmonized in total. \noxx
generated from the Energy sector provides an example of an emissions species and
sector in which all regions were satisfactorily harmonized with the default
methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
Table \ref{tab:nox} describes the parameters that underlie the choice of method
for each harmonized trajectory. Importantly, the default methods provide
regional trajectories for both the reference and mitigation scenarios closely
matching the unharmonized model behavior and also result in global aggregates
that agree with the unharmonized trajectory over the modeled time period.

@@

Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
the harmonization year for carbon monoxide (CO) emissions in the industrial
sector specifically for the \gls{sas} and \gls{afr} regions. Furthermore,
emissions in both regions see relatively large expansion or contraction,
depending on the scenario; therefore, both regions are candidates for choosing
harmonization method overrides. The default method chosen in both cases
(\code{constant_ratio}) maintains model trends for the region; however, overall
model results are distorted as can be seen in both the regional and global
panels in Figure \ref{fig:co}. By applying \code{constant_offset} overrides, the
regional trends and magnitudes are maintained. The use of harmonization
overrides also ameliorates issues seen in the harmonized global trajectory of
industrial CO (Figure \ref{fig:co}, Panel \textbf{c}). Not only does the global
trajectory with overrides more accurately match the original unharmonized model
behavior and magnitude of emissions, but the relative importance of the
underlying regional trajectories is also maintained.

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized (without markers) and unharmonized (with
      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
      lines) scenarios. Panel \textbf{a} shows harmonized and
      overridden-harmonized (respectively) trajectories for the \gls{sas} region
      while panels \textbf{b} and \textbf{c} show similar trajectories for the
      \gls{afr} region and (total) global emissions, respectively. Notably, the
      \gls{sas} and \gls{afr} regional trajectory displays a distorted
      trajectory due to the harmonization-year difference between history and
      model results in both scenarios. The distortion is large enough to affect
      global results, as shown in Panel \textbf{c}.  }
  \end{center}
\end{figure}

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows an example case of emissions
trajectories for ammonia (\nht) from the agriculture sector. Again, the
\gls{sas} region shows a large discrepancy in the harmonization year ($>$150\%
in this case). The resulting trajectory harmonized with the default method
(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
reference scenario. Notably, the SSP2-45 scenario is not affected to the same
degree. While this distortion changes the magnitude of the \gls{sas} trajectory,
it largely affects the post-2080 shape of the global trajectory (Figure
\ref{fig:nh3}, Panel \textbf{b}) as well as the relative regional contributions
to the global aggregate trajectory. For example, in the original model result,
\gls{sas} \nht~ agricultural emissions contribute $\sim$30\% of total global
emissions, whereas in the unharmonized case with default methods, \gls{sas}
comprises $\sim$50\% of global emissions by 2100. By using a
\code{constant_offset} method as an override, this distortion is addressed and
more accurately reflects unharmonized results in the \gls{sas} region, the
relative importance between regions, and global results for agricultural ammonia
emissions, each of which contributes to a better harmonization quality for the
harmonized \gls{sas} trajectory. The harmonized emissions using overrides result
in a contribution of $\sim$32\% of total global emissions by the \gls{sas}
region which aligns closely with the unharmonized model results.

@@

This work presented a novel methodology and Python implementation of automated
emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation
of the processes and methods for determining the use of harmonization methods
was provided in Section \ref{sec:meths}. \code{aneris} was able to
satisfactorily harmonized over 96\% of the 1940 individual trajectories that
were analyzed in Section \ref{sec:results}. Of the remaining trajectories,
harmonization method overrides were applied, and a discussion was provided
detailing why overrides were deemed necessary.

The automated approach drastically reduces the need for expert opinion in
determining harmonization methods for each individual combination of model
region, sector, and emissions species while still providing a justifiable
explanation for each automated choice of harmonization method based on both the
historical and future emissions trajectories. Furthermore, the automated
approach continues to scale well as models become more detailed in both the
regional and sectoral dimensions. Finally, expert opinion is still allowed to
trump the automated method as determined by the algorithm via method overrides;
however, these cases are clearly documented via the meta data provided as an
output of \code{aneris} and thus can be individually explained. This provides
not only transparency and reproducibility, but also scientific integrity in the
choice of harmonization methods.

@@

There are a variety of avenues for future improvement of both the \code{aneris}
software and underlying methodology. As with any software project, additional
users will provide use cases for more robust handling of input/output concerns
and corner cases. Further configuration parameters can also be added in the
future in order to provide overrides for all gas species in a given sector or
region. Perhaps the most fruitful investigation will involve further refinement
of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
missing from the decision tree is input from models regarding whether missing
sources or activity levels are the likely cause of a harmonization year
discrepancy (suggesting the use of an offset method) or instead a significant
difference in emissions factors (suggesting the use of a ratio method)
\cite{rogelj_discrepancies_2011}.

>>>>

\glspl{iam} are tools used to understand the complex interactions between
energy, economy, land use, water, and climate systems. \glspl{iam} provide
global projections of systemic change by dividing the world into a number of
representative regions (typically 10 to 30), the definition of which is distinct
for each model \cite{krey_global_2014}. Results from \glspl{iam} are integral in
a number of international studies, which notably include projections of climate
and energy futures. Recently, the \gls{iam} community has developed scenarios
based on the \glspl{ssp} \cite{fujimori_ssp3:_2017, van_vuuren_energy_2017,
  calvin_ssp4:_2017, fricko_marker_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The \glspl{ssp} are designed to
be used in research that include \gls{esm} simulations, climate impact,
adaptation and climate mitigation studies \cite{vuuren_new_2013}.

@@

Separately, if there are large discrepancies between the model results in the
base year and the historic data used for harmonization, convergence methods can
result in harmonized trajectories that do not faithfully represent the
underlying drivers of emissions. Furthermore, if models report negative
emissions, as is possible in scenarios designed to depict the deployment of
climate mitigation policies with large \cotwo reductions and storage,
then end-of-century emissions characteristics can be of chief import in order to
faithfully match the design parameters of the original scenarios, such as global
mean temperature and other climate metrics.

@@

A number  methods are implemented in \code{aneris} including
ratio-convergence shown in Equation \ref{eqs:ratio}, offset-convergence shown in
Equation \ref{eqs:offset}, and linear interpolation shown in Equation
\ref{eqs:interpolate}. In all equations the region, species, and sector indices
have been dropped for clarity. Each equation is a function of time, model
trajectory, historical trajectory, base year ($t_i$), and a convergence year
($t_f$), at which point the harmonized trajectory converges to the unharmonized
trajectory. \codeb{aneris} provides a number of methods to choose from for each
of the harmonization families. A summary of all available methods is provided in
Table \ref{tab:meths}.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{decision_tree.pdf}
    \caption[]{
      \label{fig:decision_tree}
      The default method decision tree used in the \code{aneris} software
      library. For all decisions, upper (purple) branches represent a ``yes''
      response and lower (orange) branches represent a ``no'' response. The
      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
      decision-making thresholds for $c_v$ and $dH$ are described in the main text.  Where
      present, convergence years of default methods are provided below the
      method name in parentheses.  Convergence years are chosen in order to
      balance the three harmonization goals discussed in Section
      \ref{sec:concept}. Methods labeled in \textit{green} are likely to closely
      match unharmonized results, methods in \textit{yellow} will likely
      somewhat match unharmonized results, and methods in \textit{red} can be
      expected to have a large relative difference between harmonized and
      unharmonized results.}
  \end{center}
\end{figure}

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is rare but possible for a sector to have emissions 
reported in the historical data, but the model to report zero for the harmonization year, 
with non-zero future values. In such cases,
an offset method is required as a ratio method would mask future emissions and
erroneously harmonize the trajectory to zero.

In most cases, however, models do report values in the harmonization
year. Figure \ref{fig:cases} displays a number of example trajectories which
highlight the possible issues resulting from harmonizing model results in
different contexts. These trajectories do not correspond to specific model
results; rather they serve as illustrative examples of the kinds of trajectories
observed in practice. Panels \code{a} and \code{b} highlight examples where
model results peak mid-century, behavior that is seen in a number of scenarios
with general emissions mitigation effects, such as pollution controls applied by
developing nations on transport and industry sectors. Panel \code{a} highlights
a case where model base-year values and history are relative close whereas Panel
\code{b} shows a situation where model values and history are relatively far
apart. Panels \code{c} and \code{d} show similar model trajectories that peak
mid-century but also have negative emissions. Models can report negative
emissions for \cotwo in future scenarios with
climate mitigation enacted via the deployment of carbon capture and
storage technologies. Again, the
relative difference with historical values differ between the panels to explore
harmonization method choices in each situation.

@@

Finally, the default harmonization decisions depend on the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
\gls{iam} values used in the \gls{ssp} and (ongoing) \gls{cmip6} inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

@@

The \texttt{Harmonizer} is a class whose task is to harmonize model value
trajectories given historical data and possible user method \textit{overrides},
i.e., non-default methods (described further in Section \ref{sec:overrides}). It
is used by the \texttt{HarmonizationDriver}; however it is also available to the
user as a first-class object. The \texttt{Harmonizer} requires that input data
conform to the \code{aneris} calculation data format, which explicitly separates
the emissions species from the sector contributing the emissions (these are
combined in the single \code{Variable} column in the standard \gls{iam}
format). Because the \texttt{Harmonizer} is designed to operate on a single
instance of a model and scenario, the canonical data format includes
\texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
without extraneous meta-data columns for the model and scenario. Once configured
with appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
    \caption[]{
      \label{fig:software}
      The various objects and their relation to one another in the \code{aneris}
      code base as well as a short description of their scope.  
    }
  \end{center}
\end{figure}

@@

%% \begin{figure}
%%   \subfloat[Python Implementation\label{fig:software}]{
%%     \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
%%     }
%%   \subfloat[Workflow\label{fig:workflow}]{
%%     \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
%%     }
%%     \caption{ Panel \ref{fig:software} shows the various objects in the
%%       \code{aneris} code base as well as a short description of their scope.
%%       Panel \ref{fig:workflow} displays the full harmonization process
%%       as executed by \code{aneris} for a single instance of a model and
%%       scenario. Operations that can be configured with user-based input
%%       configurations are shown in purple. Operations governed by the
%%       \texttt{HarmonizationDriver} are shown in red. Data processing operations
%%       are shown in blue. The core harmonization process, governed by the
%%       \texttt{Harmonizer} is shown in green.  }
%%     \label{fig:software-workflow}
%% \end{figure}

The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
by cleaning input data. Cleaning operations include adding model
trajectories with 0 values where a sector/emission combination exists in the historical data set but are not provided by the
model input and detecting any issues that would cause the harmonization process
to fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
spatial aggregation to common analysis regions is performed. For example, the
5-region aggregation developed in the \glspl{rcp}
\cite{vuuren_representative_2011} process is commonly used in the \gls{iam}
community and is shown in Figure \ref{fig:regions}. Finally, any exogenous
trajectories the user provides are added. Exogenous trajectories are normally
provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
chlorofluorocarbons provided by the \gls{wmo} \cite{wmo2014}. Upon completion,
the harmonized trajectories and meta data regarding the harmonization process
are returned. A description of all returned meta data is provided in Table
\ref{tab:metadata}.

@@

In order to help identify cases where overrides may be needed,
harmonization \textit{diagnostics} are provided which report the relative
difference between harmonized and unharmonized trajectories at their mid and
end-points when these values exceed specified threshold. The diagnostic reporting thresholds  are configurable by the user, but defaults of $400$\%
and $200$\%, respectively, are provided based on experiences of the authors'
use of \code{aneris} to date.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization results for two scenarios from
the \gls{ssp} scenario library\footnote{We refer the reader to the broad
  literature discussing the SSPs, e.g.,
  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
  depth discussion of the scenario architecture and design.} are presented here. We use the
SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
(referred to as SSP2-Ref) as an example because MESSAGE-GLOBIOM is
the marker scenario\footnote{The ``marker'' scenario concept is used to
  designate the archetype scenario used as a reference within each SSP scenario family. See
  \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
  description.} for this \gls{ssp}. This SSP2 scenario lies between two
\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing in this concept is the energy inbalance at the top of the atmsophere caused by anthropogenic influences relative to a pre-industrial reference point. Higher radiative forcing leads to larger global changes, such as surface temperture. We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
  for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
additionally present the results for the SSP2-based mitigation scenario leading
to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
scenario is chosen because mitigation technologies and policies are enacted
causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
negative \cotwo emissions in some regions and sectors due to carbon capture and
sequestration and afforestation. A scenario in which negative emissions play a
role in mitigation strategies is particularly important because of the
sensitivity of key indicators, such as end-of-century radiative forcing (which
is used to estimate mean global temperature response), to the timing and
magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
scenarios represent two contrasting cases in the use of a harmonization approach
and thus provide a case study as to its general applicability. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
aggregate \gls{ghg} emissions, in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions, and harmonization is performed at the native regional resolution of each model. Historical
data is taken from previously described \gls{luc} and anthropogenic sources,
which comprise 10 separate pollutant and \gls{ghg} species and 12 sectors shown
in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
each scenario, and therefore 1940 trajectories were harmonized in total for these two illustrative scenarios. 
\noxx
generated from the Energy sector provides an example of an emissions species and
sector in which all regions were satisfactorily harmonized with the default
methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
Table \ref{tab:nox} describes the parameters that underlie the choice of method
for each harmonized trajectory. Importantly, the default methods provide
regional trajectories for both the reference and mitigation scenarios that
match the general unharmonized model behavior and also result in global aggregated emissions
that are similar to the unharmonized trajectory.

@@

Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
the harmonization year for carbon monoxide (CO) emissions in the industrial
sector specifically for the \gls{sas} and \gls{afr} regions. Furthermore,
emissions in both regions see relatively large expansion or contraction,
depending on the scenario; therefore, both regions are candidates for choosing
harmonization method overrides. The default method chosen in both cases
(\code{constant_ratio}) maintains model trends for the region; however, overall
model results are distorted as can be seen in both the regional and global
panels in Figure \ref{fig:co}. By applying \code{constant_offset} overrides, the
regional trends and magnitudes are maintained. The use of harmonization
overrides also ameliorates issues seen in the harmonized global trajectory of
industrial CO (Figure \ref{fig:co}, Panel \textbf{c}). Not only does the global
trajectory with overrides more closely match the original unharmonized model
behavior and magnitude of emissions, but the relative importance of the
underlying regional trajectories is also maintained.

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized (without markers) and unharmonized (with
      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
      lines) scenarios. The left column shows the deault harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
      while panels \textbf{b} and \textbf{c} show  trajectories for the
      \gls{afr} region and (total) global emissions, respectively. Notably, the
      \gls{sas} and \gls{afr} regional trajectories are distorted
      when the default methods are used due to the harmonization-year difference between history and
      model results in both scenarios. The distortion is large enough to affect
      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistantcy with the unharmonized model scenario at the regional and global levels. }
  \end{center}
\end{figure}

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows an example case of emissions
trajectories for ammonia (\nht) from the agriculture sector. Again, the
\gls{sas} region shows a large discrepancy in the harmonization year ($>$150\%
in this case). The resulting trajectory harmonized with the default method
(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
reference scenario. Notably, the SSP2-45 scenario is not affected to the same
degree. While this distortion changes the magnitude of the \gls{sas} trajectory,
it largely affects the post-2080 shape of the global trajectory (Figure
\ref{fig:nh3}, Panel \textbf{b}) as well as the relative regional contributions
to the global aggregate trajectory. For example, in the original model result,
\gls{sas} \nht~ agricultural emissions contribute $\sim$30\% of total global
emissions, whereas in the harmonized case with default methods, \gls{sas}
comprises $\sim$50\% of global emissions by 2100. By using a
\code{constant_offset} method as an override, this distortion is addressed and
more accurately reflects unharmonized results in the \gls{sas} region, the
relative importance between regions, and global results for agricultural ammonia
emissions, each of which contributes to a better harmonization quality for the
harmonized \gls{sas} trajectory. The harmonized emissions using overrides result
in a contribution of $\sim$32\% of total global emissions by the \gls{sas}
region which aligns closely with the unharmonized model results.

@@

This work presented a novel methodology and Python implementation of automated
emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation
of the processes and methods for determining the use of harmonization methods
was provided in Section \ref{sec:meths}. \code{aneris} was able to
satisfactorily harmonized over 96\% of the 1940 individual trajectories that
were analyzed in Section \ref{sec:results}. Of the remaining trajectories,
harmonization method overrides were applied, and  discussion was provided
detailing why overrides were deemed necessary.

The automated approach drastically reduces the need for expert opinion in
determining harmonization methods for each individual combination of model
region, sector, and emissions species while still providing a justifiable
explanation for each automated choice of harmonization method based on both the
historical and future emissions trajectories. Furthermore, the automated
approach should continue to scale well as models become more detailed in both the
regional and sectoral dimensions. Finally, expert opinion is still allowed to
trump the automated method as determined by the algorithm via method overrides;
however, these cases are clearly documented via the meta data provided as an
output of \code{aneris} and thus can be individually explained. This provides
not only transparency and reproducibility, but also scientific integrity in the
choice of harmonization methods.

@@

There are a variety of avenues for future improvement of both the \code{aneris}
software and underlying methodology. As with any software project, additional
users will provide use cases for more robust handling of input/output issues
and corner cases. Further configuration parameters can also be added in the
future in order to provide overrides for all gas species in a given sector or
region. Perhaps the most fruitful investigation will involve further refinement
of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
missing from the decision tree is input from models regarding whether missing
sources or activity levels are the likely cause of a harmonization year
discrepancy (suggesting the use of an offset method) or instead a significant
difference in emissions factors (suggesting the use of a ratio method)
\cite{rogelj_discrepancies_2011}.
